Precision calculations of the cosmic shear power spectrum projection

### 2017-02-15 conclusions.tex
Peter's comments IV

\mk{TODO: comment on non-flat models.}

>>>>

\mk{ We have considered a flat Universe throughout this paper. To generalise
the calculations to non-flat models, one needs to modify the comoving angular
diameter distance to account for the spatial curvature $K \ne 0$. In addition,
the spherical Bessel functions are replaced with hyperspherical Bessel
functions \citep{1986ApJ...308..546A}. In a universe with positive curvature
$K>0$ the 3D wave modes become discrete integer variables. For non-flat models,
we do however not expect qualitative differences from our results.}

### 2017-02-15 limber2.tex
Peter's comments IV

\newcommand{\mk}[1]{{\bf\textcolor{blue}{#1}}}
\newcommand{\Label}[1]{\quad \mk{(\mbox{#1})}}
\newcommand{\LabelTxt}[1]{\mk{(\mbox{#1})}}

@@

The lensing potential $\psi$ at a direction (\MA {position}?) on the sky $(\theta, \varphi)$ in
the Born approximation is defined as the projected 3D metric potential $\Phi$
along the line of sight of a flat Universe \citep{1998ApJ...498...26K,BS01},
%
% Hu00 (21,22) with minus sign
%
\begin{equation}
  \psi(\theta, \varphi) = \frac 2 {{\rm c}^2} \int\limits_0^\infty \frac{{\rm d}\chi}{\chi}
    %\Phi[\chi, f_K(\chi) \theta, f_K(\chi) \varphi; \chi] \, q(\chi),
    \Phi[\chi, \chi \theta, \chi \varphi; \chi] \, q(\chi),
  \label{eq:psi}
\end{equation}
%
where the lensing efficiency $q$ is given by
%
\begin{equation}
  q(\chi) = \int\limits_\chi^{\chi_{\rm h}} {\rm d} \chi^\prime \, n(\chi^\prime)
    %\frac{f_K(\chi^\prime - \chi)}{f_K(\chi^\prime)},
    \frac{\chi^\prime - \chi}{\chi^\prime},
  %
  \label{eq:lens_efficiency}
\end{equation}
%
corresponding to a population of lensed galaxies with a normalised source
redshift distribution $n_z(z) {\rm d}z = n(\chi) {\rm d} \chi$, with the limit
being the comoving distance to the horizon $\chi_{\rm h}$\footnote{\mk{In
(\ref{eq:psi}) we have replaced this limit without loss of generality by
$\infty$, since $q(\chi > \chi_{\rm h}) = 0$.}}. Here, ${\rm c}$ is the speed of
light, and the projection is carried out over comoving distances $\chi$
%, and $f_K(\chi) = |K|^{-1/2} S_K(|K|^{1/2} \chi)$ is the comoving angular
% diameter distance, where the function $S_K$ equals $\sin$, ${\rm id}$, or
% $\sinh$ if the spatial curvature of the Universe $K$ is positive, zero, or
% negative, respectively.
The last argument of the potential $\Phi$ is not to
be understood as coordinate, but rather as a substitute of cosmic time,
$t(\chi)$, to express the time at which the potential is evaluated. This is
true in the following for all fields and functions thereof that dynamically
change with cosmic epoch.

@@

Following \cite{2000PhRvD..62d3007H} we decompose the lensing
potential (equation~\ref{eq:psi}) into spherical harmonics, in analogy
to the \MA{cosmic microwave background (CMB)}
temperature, both of which are scalar functions on the sphere. This
decomposition and its inverse are
%
% Hu00 (23)
%
\begin{align}
  \psi(\theta, \varphi) = \sum_{\ell=0}^\infty \sum_{m=-\ell}^\ell \psi_{\ell m} {\rm Y}_{\ell m}(\theta, \varphi);
    \label{eq:psi_harm_exp}
    \\
    %
  \psi_{\ell m} = \int_{\mathbb{S}^2} {\rm d} \Omega \, \psi(\theta, \varphi) {\rm Y}^\ast_{\ell m}(\theta, \varphi).
  \label{eq:psi_harm_exp_inv}
\end{align}
%
Complex conjugation is denoted with the superscript $^\ast$.
To specify tomographic redshift bins $i=0\ldots N_z-1$, we introduce a family
of lensing efficiency functions $q_i$ defined by a corresponding family of
redshift distributions $n_i$ via equation~(\ref{eq:lens_efficiency}). The resulting
lensing potential is denoted by $\psi_{\ell m, i}$. The tomographic power
spectrum of the lensing potential between two redshift bins $i$ and $j$,
$C_{ij}^\psi$ \citep{pee80} is then defined as
%
\begin{equation}
  \left\langle \psi^{}_{\ell m, i} \, \psi^\ast_{\ell^\prime m^\prime, j} \right\rangle
    = \delta_{\ell \ell^\prime} \delta_{m m^\prime} C^\psi_{ij}(\ell) .
  \label{eq:C_ell_psi}
\end{equation}
%
\mk{Note that the argument $\ell$ is a discrete integer variable, and is often written as index, $C_\ell$.}
Using the properties of the spherical harmonics (see
App.~\ref{sec:derivations_C} for details) the power spectrum can be written as
%
\begin{align}
  C_{ij}^\psi(\ell) = & \frac 8 {c^4 \pi} 
  \int_0^\infty \frac{{\rm d}\chi}{\chi} q_i(\chi)
  \int_0^\infty \frac{{\rm d}\chi^\prime}{\chi^\prime} q_j(\chi^\prime)
  \int {\rm d} k k^2 \, {\rm j}_\ell(k \chi) {\rm j}_\ell(k \chi^\prime) P_\Phi(k; \chi, \chi^\prime)
  \label{eq:C_ell_phi_Pphi} \\
  = & \frac 8 \pi \pref^2
  \int_0^\infty \frac{{\rm d}\chi}{\chi} \frac{q_i(\chi)}{a(\chi)}
  \int_0^\infty \frac{{\rm d}\chi^\prime}{\chi^\prime} \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
  \int \frac{{\rm d} k}{k^2} \, {\rm j}_\ell(k \chi) {\rm j}_\ell(k \chi^\prime) P_{\rm m}(k; \chi, \chi^\prime) ;
  \label{eq:C_ell_phi_Pm}
\end{align}
%
where ${\rm j}_\ell$ is the spherical Bessel function of order $\ell$. For
convenience we introduce the normalisation constant $\pref$ as
%
\begin{equation}
  \pref = \frac 3 2 \Omega_{\rm m} \left(\frac{H_0} c\right)^2.
  \label{eq:pref}
\end{equation}
%
The first line expresses $C_{ij}^\psi$ in terms of the 3D potential power spectrum $P_\Phi$, defined as
%
\begin{align}
  \left\langle \hat \Phi(\vec k; \chi) \hat \Phi^\ast(\vec k^\prime; \chi^\prime) \right\rangle
    = & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) P_\Phi(k; \chi, \chi^\prime).
    %\nonumber \\
    %= & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) \pref^2 k^{-4} a^{-1}(\chi) a^{-1}(\chi^\prime)
      %P_{\rm m}(k; \chi, \chi^\prime).
  \label{eq:p_phi}
\end{align}
%
The second line uses the 3D matter density power spectrum $P_{\rm m}$, which is defined analogously as
%
\begin{align}
  \left\langle \hat \delta(\vec k; \chi) \hat \delta^\ast(\vec k^\prime; \chi^\prime) \right\rangle
    = & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) P_{\rm m}(k; \chi, \chi^\prime),
  \label{eq:p_m}
\end{align}
%
and is related to $P_\Phi$ via the absolute square of the Poisson equation
(\ref{eq:poisson}).

@@

The second-order Limber term in equation~(\ref{eq:C_ell_limber12_dr}) has an additional
$\nu^{-2}$-dependence, and is therefore strongly suppressed for large $\ell$.
%
\begin{align}
  C^\gamma_{{\rm L2}, ij}(\ell) = & - \frac 1 {\nu^2(\ell)} \, \frac{\ellbar^2(\ell, 2)}{\nu^4(\ell)} \, \pref^2
    \int {\rm d} \chi \, \chi^{2} a^{-2} P^{1/2}_{\rm m}\left(\frac{\nu(\ell)}\chi; \chi\right)
    %\nonumber \\
    %& \times
    \frac{a(\chi)}{2} \chi^{3/2} \left[ q^{}_i f^{\prime\prime}_j + f^{\prime\prime}_i q^{}_j
      + \frac{\chi}{3} \left( q^{}_i f^{\prime\prime\prime}_j + f^{\prime\prime\prime}_i q^{}_j
      \right)
    \right](\nu(\ell)/\chi, \chi) \, . 
  \label{eq:C_ell_limber2_dr} 
\end{align}
%
The higher-order derivatives of the filter functions have to be computed
numerically in the general case. These suffer from numerical noise and are
sensitive to the set-up, for example the step size. The tabulation and
interpolation of those derivatives is time-taking since they depend on two
arguments, $\nu$ and $\chi$. In the following section, we will separate the
$k$- and $\chi$-dependent parts of the power spectrum to make the numerical
derivatives faster and more smooth.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsubsection{Time-independent power spectrum separation}
\subsubsection{Facoring out the growth factor}
\label{sec:factor_D}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\mk{TODO: Make this clearer.}

With the definition (\ref{eq:P_k_chi_sep_s}), we factor out the function
$\tilde P_{\rm m}(k, \chi)$ from equation~(\ref{eq:f_LA08}), and we define the
separated kernel function $f_{\rm s}$ as
%
\begin{equation}
  f_{\rm s}(\chi) = D_+(\chi) a^{-1}(\chi) \chi^{-3/2} q(\chi) \,
  \label{eq:f_LA08_s}
\end{equation}
%
The second-order Limber term of the shear power spectrum can then be approximated as
%
\begin{align}
  C^\gamma_{{\rm L2}, ij}(\ell) = & - \frac 1 {\nu^2(\ell)} \, \frac{\ellbar^2(\ell, 2)}{\nu^4(\ell)} \, \pref^2
    \int {\rm d} \chi \, \chi^{2} a^{-2} P_{\rm m}\left(\frac{\nu(\ell)}\chi; \chi\right)
    %\nonumber \\
    %& \times
    \frac {a(\chi)}{2 D_+(\chi)} \chi^{3/2}
    \left[ q^{}_i f^{\prime\prime}_{{\rm s} j} + f^{\prime\prime}_{{\rm s}i} q^{}_j
      + \frac{\chi}{3} \left( q^{}_i f^{\prime\prime\prime}_{{\rm s}j} + f^{\prime\prime\prime}_{{\rm s}i} q^{}_j
      \right)
    \right](\chi) \, . \Label{ExtL2Sph}
  \label{eq:C_ell_limber2_dr_s}
\end{align}
%
We compute the numerical higher derivatives as follows. The functions $f_{\rm
s}(\chi)$ are fitted as power laws with index $\approx 1.5$. The fit is
performed between $\chi_{\rm min} = 0.001 {\rm Mpc}/h$ and $\chi_{\rm max} =
500 \mbox{Mpc}/h$. At larger comoving distances the kernel decreases faster
than a power law, so we exclude this range from the fit. Even though the
derivaties are larger due to the steeper decline, the associated errors are
very small as these scales are down-weighed by the kernel function $f_{\rm
s}$ itself. At $\chi = 500 \, \mbox{Mpc}/h$ the filter function is less than
$10^{-4}$ of its value of unity at $1$\,Mpc$/h$.

@@

The extended flat-sky Limber approximation is readily derived from the
spherical case (\ref{eq:C_ell_limber1}, \ref{eq:C_ell_limber2_dr_s}), by
replacing the pre-factor $\ellbar^2(\ell, 2)$ with $\ell^4$,
%
\begin{align}
  P^\gamma_{{\rm L1}, ij}(\ell) = & \, p(\ell) \, \pref^2 \int {\rm d} \chi \frac{ q_i(\chi) q_j(\chi) }{a^2(\chi)}
  P_{\rm m}\left(\frac{\nu(\ell)}{\chi}; \chi\right);
  \label{eq:P_ell_limber1}
  \\
    P^\gamma_{{\rm L2}, ij}(\ell) = & - \frac 1 {\nu^2(\ell)} \, p(\ell) \, \pref^2
    \int {\rm d} \chi \, \chi^{2} a^{-2} P_{\rm m}\left(\frac{\nu(\ell)}\chi; \chi\right)
    %\nonumber \\
    %& \times
    \frac{a(\chi)}{2 D_+(\chi)} \chi^{3/2} \left[ q^{}_i f^{\prime\prime}_{{\rm s}j} + f^{\prime\prime}_{{\rm s}i} q^{}_j  
      + \frac{\chi}{3} \left( q^{}_i f^{\prime\prime\prime}_{{\rm s}j} + f^{\prime\prime\prime}_{{\rm s}i} q^{}_j
      \right)
    \right](\chi)
  \label{eq:P_ell_limber2}
\end{align}
%
Further approximations are can made for the pre-factor $p(\ell) = \ell^4/\nu^4(\ell)$ and $\nu(\ell)$:
%
\begin{enumerate}
  \item $p(\ell) = 1$, this corresponds to $\nu(\ell) = \ell$, which is the
    \emph{standard} Limber approximation \LabelTxt{L1Fl}
    Until recently, i.e.~for all pre-2014 CFHTLenS results, this was the approximation of choice.
    Note that we do not discuss the non-extended second-order Limber approximation here.
  %
  \item $p(\ell) = \ell^4/(\ell + 1/2)^4$. This corresponds to the \emph{extended} Limber
    approximation \LabelTxt{ExtL1Fl,ExtL2Fl} with $\nu(\ell) = \ell + 1/2$; however the following case is typically employed:
  %
  \item $p(\ell) = 1$, but keeping as argument of the power spectrum $\nu(\ell) = \ell + 1/2$. This is
    a \emph{hybrid} between
    standard and extended Limber approximation \Label{ExtL1FlHyb,ExtL2FlHyb}, and this case of first order
    was used in \cite{KiDS-450,joudaki/etal:2016}. As is shown
    later, this is a better approximation to the full projection than case (ii). In (\ref{eq:P_ell_limber2})
    the second-order suppression factor is also left to be $\nu^{-2}(\ell) = (\ell + 1/2)^{-2}$, providing
    a slightly more accurate approximation compared to $\nu^{-2}(\ell) = \ell^{-2}$.
\end{enumerate}
%

@@

Compared to the statistical power of current and future surveys, all
approximations discussed here are subdominant to the cosmic variance, $\Delta
C(\ell) / C(\ell) = [f_{\rm sky} (2 \ell + 1)]^{-1/2}$
\citep{1992ApJ...388..272K}, where $f_{\rm sky}$ is the faction of sky observed
by the survey. The uncertainties from the Limber approximation in the case of
ExtL2Sph is an order of magnitude below the cosmic variance of a Euclid-like
survey (sky area $15,000$ square degrees) for all $\ell$.

@@

Inserting equation~(\ref{eqn:B1}) into the second-order Limber power spectrum
in equation~(\ref{eq:C_ell_limber2_dr_s}) and using the inverse Poisson
equation to replace the matter with the potential power spectrum we obtain the
same expression as \cite{vande2012} (their Equation~7.19), except for an
additional factor $a^2(\chi)$. This difference is accounted for by the factor
$a^{-1}$ in the propagation of primordial potential perturbations to late
times, see Equation~(6.22) in \cite{vande2012}, where the transition is
calculated using $g(a)$ instead of $D(a)$. \mk{TODO: Check this further.}

>>>>

\newcommand{\mk}[1]{{\bf\textcolor{blue}{#1}}}
\newcommand{\Label}[1]{\quad (\mbox{#1})}
\newcommand{\LabelTxt}[1]{(\mbox{#1})}

@@

The lensing potential $\psi$ at a direction (\MA {position}?) on the sky $(\theta, \varphi)$ in
the Born approximation is defined as the projected 3D metric potential $\Phi$
along the line of sight of a flat Universe \citep{1998ApJ...498...26K,BS01},
%
% Hu00 (21,22) with minus sign
%
\begin{equation}
  \psi(\theta, \varphi) = \frac 2 {{\rm c}^2} \int\limits_0^\infty \frac{{\rm d}\chi}{\chi}
    %\Phi[\chi, f_K(\chi) \theta, f_K(\chi) \varphi; \chi] \, q(\chi),
    \Phi[\chi, \chi \theta, \chi \varphi; \chi] \, q(\chi),
  \label{eq:psi}
\end{equation}
%
where the lensing efficiency $q$ is given by
%
\begin{equation}
  q(\chi) = \int\limits_\chi^{\chi_{\rm h}} {\rm d} \chi^\prime \, n(\chi^\prime)
    %\frac{f_K(\chi^\prime - \chi)}{f_K(\chi^\prime)},
    \frac{\chi^\prime - \chi}{\chi^\prime},
  %
  \label{eq:lens_efficiency}
\end{equation}
%
corresponding to a population of lensed galaxies with a normalised source
redshift distribution $n_z(z) {\rm d}z = n(\chi) {\rm d} \chi$, with the limit
being the comoving distance to the horizon $\chi_{\rm h}$\footnote{\mk{In
(\ref{eq:psi}) we have replaced this limit without loss of generality by
$\infty$, since $q(\chi > \chi_{\rm h}) = 0$.}}. Here, ${\rm c}$ is the speed of
light, and the projection is carried out over comoving distances $\chi$.
%, and $f_K(\chi) = |K|^{-1/2} S_K(|K|^{1/2} \chi)$ is the comoving angular
% diameter distance, where the function $S_K$ equals $\sin$, ${\rm id}$, or
% $\sinh$ if the spatial curvature of the Universe $K$ is positive, zero, or
% negative, respectively.
The last argument of the potential $\Phi$ is not to
be understood as coordinate, but rather as a substitute of cosmic time,
$t(\chi)$, to express the time at which the potential is evaluated. This is
true in the following for all fields and functions thereof that dynamically
change with cosmic epoch.

@@

Following \cite{2000PhRvD..62d3007H} we decompose the lensing
potential (equation~\ref{eq:psi}) into spherical harmonics, in analogy
to the \MA{cosmic microwave background (CMB)}
temperature, both of which are scalar functions on the sphere. This
decomposition and its inverse are
%
% Hu00 (23)
%
\begin{align}
  \psi(\theta, \varphi) = \sum_{\ell=0}^\infty \sum_{m=-\ell}^\ell \psi_{\ell m} {\rm Y}_{\ell m}(\theta, \varphi);
    \label{eq:psi_harm_exp}
    \\
    %
  \psi_{\ell m} = \int_{\mathbb{S}^2} {\rm d} \Omega \, \psi(\theta, \varphi) {\rm Y}^\ast_{\ell m}(\theta, \varphi).
  \label{eq:psi_harm_exp_inv}
\end{align}
%
Complex conjugation is denoted with the superscript $^\ast$.
To specify tomographic redshift bins $i=1\ldots N_z$, we introduce a family
of lensing efficiency functions $q_i$ defined by a corresponding family of
redshift distributions $n_i$ via equation~(\ref{eq:lens_efficiency}). The resulting
lensing potential is denoted by $\psi_{\ell m, i}$. The tomographic power
spectrum of the lensing potential between two redshift bins $i$ and $j$,
$C_{ij}^\psi$ \citep{pee80} is then defined as
%
\begin{equation}
  \left\langle \psi^{}_{\ell m, i} \, \psi^\ast_{\ell^\prime m^\prime, j} \right\rangle
    = \delta_{\ell \ell^\prime} \delta_{m m^\prime} C^\psi_{ij}(\ell) .
  \label{eq:C_ell_psi}
\end{equation}
%
\mk{Note that the argument $\ell$ is a discrete integer variable, and is often written as index, $C_\ell$.}
Using the properties of the spherical harmonics (see
App.~\ref{sec:derivations_C} for details) the power spectrum can be written as
%
\begin{align}
  C_{ij}^\psi(\ell) = & \frac 8 {c^4 \pi} 
  \int_0^\infty \frac{{\rm d}\chi}{\chi} q_i(\chi)
  \int_0^\infty \frac{{\rm d}\chi^\prime}{\chi^\prime} q_j(\chi^\prime)
  \int {\rm d} k k^2 \, {\rm j}_\ell(k \chi) {\rm j}_\ell(k \chi^\prime) P_\Phi(k; \chi, \chi^\prime)
  \label{eq:C_ell_phi_Pphi} \\
  = & \frac 8 \pi \pref^2
  \int_0^\infty \frac{{\rm d}\chi}{\chi} \frac{q_i(\chi)}{a(\chi)}
  \int_0^\infty \frac{{\rm d}\chi^\prime}{\chi^\prime} \frac{q_j(\chi^\prime)}{a(\chi^\prime)}
  \int \frac{{\rm d} k}{k^2} \, {\rm j}_\ell(k \chi) {\rm j}_\ell(k \chi^\prime) P_{\rm m}(k; \chi, \chi^\prime) ;
  \label{eq:C_ell_phi_Pm}
\end{align}
%
where ${\rm j}_\ell$ is the spherical Bessel function of order $\ell$. For
convenience we introduce the normalisation constant $\pref$ as
%
\begin{equation}
  \pref = \frac 3 2 \Omega_{\rm m} \left(\frac{H_0} c\right)^2.
  \label{eq:pref}
\end{equation}
%
The first line expresses $C_{ij}^\psi$ in terms of the 3D potential power spectrum $P_\Phi$, defined as
%
\begin{align}
  \left\langle \hat \Phi(\vec k; \chi) \hat \Phi^\ast(\vec k^\prime; \chi^\prime) \right\rangle
    = & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) P_\Phi(k; \chi, \chi^\prime).
    %\nonumber \\
    %= & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) \pref^2 k^{-4} a^{-1}(\chi) a^{-1}(\chi^\prime)
      %P_{\rm m}(k; \chi, \chi^\prime).
  \label{eq:p_phi}
\end{align}
%
The second line uses the 3D matter density power spectrum $P_{\rm m}$, which is defined analogously as
%
\begin{align}
  \left\langle \hat \delta(\vec k; \chi) \hat \delta^\ast(\vec k^\prime; \chi^\prime) \right\rangle
    = & (2\pi)^3 \delta_{\rm D}(\vec k - \vec k^\prime) P_{\rm m}(k; \chi, \chi^\prime),
  \label{eq:p_m}
\end{align}
%
and is related to $P_\Phi$ via the absolute square of the Poisson equation
(\ref{eq:poisson}).

@@

The second-order Limber term in equation~(\ref{eq:C_ell_limber12_dr}) has an additional
$\nu^{-2}$-dependence, and is therefore strongly suppressed for large $\ell$,
%
\begin{align}
  C^\gamma_{{\rm L2}, ij}(\ell) = & - \frac 1 {\nu^2(\ell)} \, \frac{\ellbar^2(\ell, 2)}{\nu^4(\ell)} \,
    \frac{\pref^2}{2}
    \int {\rm d} \chi \, \chi^{7/2} a^{-1} P^{1/2}_{\rm m}\left(\frac{\nu(\ell)}\chi; \chi\right)
    %\nonumber \\
    %& \times
    \left[ q^{}_i f^{\prime\prime}_j + f^{\prime\prime}_i q^{}_j
      + \frac{\chi}{3} \left( q^{}_i f^{\prime\prime\prime}_j + f^{\prime\prime\prime}_i q^{}_j
      \right)
    \right](\nu(\ell)/\chi, \chi) \, . 
  \label{eq:C_ell_limber2_dr} 
\end{align}
%
The higher-order derivatives of the filter functions have to be computed
numerically in the general case. These suffer from numerical noise and are
sensitive to the set-up, for example the step size. The tabulation and
interpolation of those derivatives is time-consuming since they depend on two
arguments, $\nu$ and $\chi$. In the following section, we will separate the
$k$- and $\chi$-dependent parts of the power spectrum to make the numerical
derivatives faster and more smooth.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{\mk{Approximation for small $\ell$}}
\label{sec:factor_D}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%\mk{TODO: Make this clearer.}

With the definition (\ref{eq:P_k_chi_sep_s}), we factor out the function
$\tilde P_{\rm m}(k, \chi)$ from equation~(\ref{eq:f_LA08}), and we define the
separated kernel function $f_{\rm s}$ as
%
\begin{equation}
  f_{\rm s}(\chi) = D_+(\chi) a^{-1}(\chi) \chi^{-3/2} q(\chi) \,
  \label{eq:f_LA08_s}
\end{equation}
%
The second-order Limber term of the shear power spectrum can then be approximated as
%
\begin{align}
  C^\gamma_{{\rm L2}, ij}(\ell) = & - \frac 1 {\nu^2(\ell)} \, \frac{\ellbar^2(\ell, 2)}{\nu^4(\ell)} \,
    \frac{\pref^2}{2}
    \int {\rm d} \chi \, \chi^{7/2} a^{-1}(\chi) D_+^{-1}(\chi) P_{\rm m}\left(\frac{\nu(\ell)}\chi; \chi\right)
    %\nonumber \\
    %& \times
    \left[ q^{}_i f^{\prime\prime}_{{\rm s} j} + f^{\prime\prime}_{{\rm s}i} q^{}_j
      + \frac{\chi}{3} \left( q^{}_i f^{\prime\prime\prime}_{{\rm s}j} + f^{\prime\prime\prime}_{{\rm s}i} q^{}_j
      \right)
    \right](\chi) \, . \Label{ExtL2Sph}
  \label{eq:C_ell_limber2_dr_s}
\end{align}
%
We compute the numerical higher derivatives as follows. The functions $f_{\rm
s}(\chi)$ are fitted as power laws with index $\approx \mk{-1.5}$, \mk{which
follows since the growth suppression factor $D_+(a)/a$ varies slowly with
$\chi$, and the lensing efficiency $q ~ 1$ for small and medium $\chi$ given
the CFHTLenS redshift range.} The fit is performed between $\chi_{\rm min} =
0.001 {\rm Mpc}/h$ and $\chi_{\rm max} = 500 \mbox{Mpc}/h$. At larger comoving
distances the kernel decreases faster than a power law, so we exclude this
range from the fit. Even though the derivaties are larger due to the steeper
decline, the associated errors are very small as these scales are down-weighed
by the kernel function $f_{\rm s}$ itself. At $\chi = 500 \, \mbox{Mpc}/h$ the
filter function is less than $10^{-4}$ of its value of unity at $1$\,Mpc$/h$.

@@

The extended flat-sky Limber approximation is readily derived from the
spherical case (\ref{eq:C_ell_limber1}, \ref{eq:C_ell_limber2_dr_s}), by
replacing the pre-factor $\ellbar^2(\ell, 2)$ with $\ell^4$,
%
\begin{align}
  P^\gamma_{{\rm L1}, ij}(\ell) = & \, p(\ell) \, \pref^2 \int {\rm d} \chi \frac{ q_i(\chi) q_j(\chi) }{a^2(\chi)}
  P_{\rm m}\left(\frac{\nu(\ell)}{\chi}; \chi\right);
  \label{eq:P_ell_limber1}
  \\
    P^\gamma_{{\rm L2}, ij}(\ell) = & - \frac 1 {\nu^2(\ell)} \, p(\ell) \, \frac{\pref^2}{2}
    \int {\rm d} \chi \, \chi^{7/2} a^{-1}(\chi) D_+^{-1}(\chi) P_{\rm m}\left(\frac{\nu(\ell)}\chi; \chi\right)
    %\nonumber \\
    %& \times
    \left[ q^{}_i f^{\prime\prime}_{{\rm s}j} + f^{\prime\prime}_{{\rm s}i} q^{}_j  
      + \frac{\chi}{3} \left( q^{}_i f^{\prime\prime\prime}_{{\rm s}j} + f^{\prime\prime\prime}_{{\rm s}i} q^{}_j
      \right)
    \right](\chi)
  \label{eq:P_ell_limber2}
\end{align}
%
Further approximations can be made for the pre-factor $p(\ell) = \ell^4/\nu^4(\ell)$ and $\nu(\ell)$:
%
\begin{enumerate}
  \item $p(\ell) = 1$, this corresponds to $\nu(\ell) = \ell$, which is the
    \emph{standard} Limber approximation \LabelTxt{L1Fl}
    Until recently, i.e.~for all pre-2014 CFHTLenS results, this was the approximation of choice.
    Note that we do not discuss the second-order Limber approximation \mk{with this $p$} here.
  %
  \item $p(\ell) = \ell^4/(\ell + 1/2)^4$. This corresponds to the \emph{extended} Limber
    approximation \LabelTxt{ExtL1Fl,ExtL2Fl} with $\nu(\ell) = \ell + 1/2$; however the following case is typically employed:
  %
  \item $p(\ell) = 1$, but keeping as argument of the power spectrum $\nu(\ell) = \ell + 1/2$. This is
    a \emph{hybrid} between
    standard and extended Limber approximation \Label{ExtL1FlHyb,ExtL2FlHyb}, and this case of first order
    was used in \cite{KiDS-450,joudaki/etal:2016}. As is shown
    later, this is a better approximation to the full projection than case (ii). In (\ref{eq:P_ell_limber2})
    the second-order suppression factor is also left to be $\nu^{-2}(\ell) = (\ell + 1/2)^{-2}$, providing
    a slightly more accurate approximation compared to $\nu^{-2}(\ell) = \ell^{-2}$.
\end{enumerate}
%

@@

Compared to the statistical power of current and future surveys, all
approximations discussed here are subdominant to the cosmic variance, $\Delta
C(\ell) / C(\ell) = [f_{\rm sky} (2 \ell + 1)]^{-1/2}$
\citep{1992ApJ...388..272K}, where $f_{\rm sky}$ is the fraction of sky observed
by the survey. The uncertainties from the Limber approximation in the case of
ExtL2Sph is an order of magnitude below the cosmic variance of a Euclid-like
survey (sky area $15,000$ square degrees) for all $\ell$.

@@

Inserting equation~(\ref{eqn:B1}) into the second-order Limber power spectrum
in equation~(\ref{eq:C_ell_limber2_dr_s}) and using the inverse Poisson
equation to replace the matter with the potential power spectrum we obtain the
same expression as \cite{vande2012} (their Equation~7.19), except for an
additional factor $a^2(\chi)$. This difference is accounted for by the factor
$a^{-1}$ in the propagation of primordial potential perturbations to late
times, see Equation~(6.22) in \cite{vande2012}, where the transition is
calculated using $g(a)$ instead of $D(a)$.
%\mk{TODO: Check this further.}
